%
\documentclass[a4paper, 10pt, conference]{ieeeconf}

\IEEEoverridecommandlockouts{}
\overrideIEEEmargins{}

\usepackage{graphics, epsfig, url}
\usepackage{amsmath, amssymb, mathptmx}
\usepackage[colorlinks=true]{hyperref}

\title{
    \LARGE \bf
    Task-parameterised Gaussian Processes
    \\ [0.5em]
    \large MTech Project MidTerm Report
}
\author{
    Ashrith Sagar Yedlapalli$^{1}$
    \thanks{$^{1}$ MTech RAS, IISc Bangalore,
    {\tt\small ashrithy@iisc.ac.in}}
}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

\begin{abstract}
    Robot learning from demonstrations (LfD) is a promising approach for teaching complex skills to robots without explicit programming.
    However, generalizing these skills to novel situations remains a challenge.
\end{abstract}

\section{INTRODUCTION}

Robot learning from demonstrations (LfD) has emerged as a powerful paradigm for teaching robots complex tasks without explicit programming.
In interactive LfD, non-expert users can kinesthetically guide the robot to perform tasks like pick-and-place, dressing, or cleaning.
However, a major challenge is generalizing these learned policies to novel situations, such as varying object positions, arm postures, or surface geometries.

To address generalization, task-parameterized models have been developed to adapt skills based on external parameters, such as reference frames representing objects or goals.
For instance, Task-Parameterized Gaussian Mixture Models (TP-GMM) encode demonstrations in multiple frames and combine them probabilistically for new configurations.
Building on this, Task-Parameterized Gaussian Processes (TPGP)~\cite{ramirez2024learning} leverage Gaussian Processes (GPs) for modeling local policies relative to each frame, with self-supervised learning of frame relevance using uncertainty quantification to improve stability and performance in multi-frame tasks, as demonstrated in simulation and robotic re-shelving experiments.

Recent advancements further enhance adaptability; for example, generalizable motion policies through keypoint parameterization~\cite{lee2024generalizable} enable robust skill transfer across varying environments by parameterizing motions with keypoints and transportation maps. Similarly, Elastic Motion Policies (EMP)~\cite{li2025elastic} provide a dynamical system for one-shot imitation learning, allowing real-time adaptation via Laplacian editing and Lyapunov functions, ensuring convergence in dynamic settings.

Interactive imitation learning algorithms, such as DAgger (Dataset Aggregation), iteratively collect expert corrections to mitigate distribution shift in behavioral cloning.
Variants like HG-DAgger (Human-Gated DAgger) incorporate human oversight to decide when interventions are needed, reducing expert burden in real-world settings.
Compliant Residual DAgger (CR-DAgger) extends this by using residual policies for compliant control, enabling safer and more adaptive learning from converging supervisors.
These methods, surveyed in~\cite{celemin2022interactive}, enhance LfD by combining demonstration with online feedback.

\section{RELATED WORK}

Movement primitives (MPs) provide compact, reusable representations of robotic motions, enabling efficient imitation learning, generalization, and adaptation.
Dynamic Movement Primitives (DMPs), introduced by Ijspeert et al.~\cite{Ijspeert2013}, model both discrete and rhythmic behaviors as autonomous dynamical systems.
A DMP consists of a canonical linear system generating a phase variable and a transformation system (second-order spring-damper) driven by a learned nonlinear forcing term approximated via locally weighted regression on basis functions.
Demonstrations are encoded into weights, allowing straightforward modulation of amplitude, goal, duration, and coupling with external forces or other primitives—making DMPs a cornerstone for trajectory generation in imitation learning.

Probabilistic Movement Primitives (ProMPs) by Paraschos et al.~\cite{10.5555/2999792.2999904} extend DMPs into a fully probabilistic framework.
Trajectories are represented as a linear combination of radial basis functions whose weights follow a multivariate Gaussian distribution learned from multiple demonstrations.
This captures movement variability, supports stochastic execution, and enables conditioning on via-points or final states through closed-form Bayesian updates.
Kernelized Movement Primitives (KMPs), proposed by Huang et al.~\cite{Huang2019}, replace fixed basis functions with kernel ridge regression, yielding a non-parametric representation.
KMPs naturally handle multiple via-point constraints, obstacle avoidance, and fusion of primitives without explicit phase alignment, offering superior flexibility for complex task adaptation in interactive imitation scenarios.

\subsection{Task-Parameterized Models}

Task-parameterized models address generalization by encoding skills relative to task parameters, such as coordinate frames.
Task-Parameterized Gaussian Mixture Models (TP-GMM), introduced by Calinon~\cite{calinon2016tutorial}, transform demonstrations into multiple reference frames and fit GMMs to each.
At execution, the models are transformed to new frame configurations and combined via Gaussian product to generate adapted trajectories.
This approach enables heuristic-free and segmentation-free learning of multi-frame skills, as shown in applications like manipulation and dressing assistance~\cite{pignat2017learning}.
However, TP-GMM relies on discrete mixtures, which may limit handling of uncertainties.

Task-Parameterized Gaussian Processes (TPGP), proposed by Ramírez Montero et al.~\cite{ramirez2024learning}, extend this framework using GPs.
Local GP policies model dynamics relative to each frame, incorporating progress variables and variance minimization for stability.
Frame relevance is learned self-supervised via another GP, optimizing likelihood on held-out demonstrations.
Evaluated on multi-frame tasks, TPGP outperforms TP-GMM in goal proximity and Fréchet distance metrics, demonstrating robust generalization in simulation and real-world robotic re-shelving.

Recent works like Lee et al.~\cite{lee2024generalizable} propose keypoint-parameterized motion policies with transportation maps for generalizable skill learning, enabling adaptation to new scenes without retraining. Elastic Motion Policies~\cite{li2025elastic} introduce a dynamical system for one-shot imitation, adapting via Laplacian editing and online Lyapunov learning for robust performance in dynamic environments.

\subsection{Interactive Imitation Learning}

Interactive imitation learning enhances standard LfD by allowing online expert feedback to correct learner errors.
DAgger (Dataset Aggregation), by Ross et al.~\cite{ross2011reduction}, iteratively trains a policy on an aggregated dataset of states visited by the learner and actions provided by the expert, addressing covariate shift in behavioral cloning.
This no-regret online learning reduction provides strong performance guarantees.

HG-DAgger (Human-Gated DAgger), introduced by Kelly et al.~\cite{kelly2019hg}, modifies DAgger for human experts by gating interventions based on estimated risk or uncertainty, reducing supervision time while maintaining safety in real-world systems like robotics.

Compliant Residual DAgger (CR-DAgger) builds on DAgger with residual policies for compliant control, enabling learning from converging supervisors~\cite{balakrishna2019imitation}.
It focuses on practical implementation in contact-rich tasks, providing guidance for effective data collection and human corrections in real-world robot learning.

These techniques are comprehensively surveyed in~\cite{celemin2022interactive}, highlighting their role in bridging demonstration and reinforcement learning for adaptive robotics.

\section{METHODOLOGY}

The proposed methodology follows from the TPGP framework for learning multi-reference frame skills from demonstrations.
The core idea is to transform demonstration data to local reference frames to encode relative dynamics, and then use a self-supervised approach to train a frame relevance predictor that selects the most relevant frame during execution. A diagram showing the main steps is illustrated in Fig. 2 of the original paper.

\subsection{Demonstration Recording}
During the recording of a demonstration, the 2D/3D position of the agent (e.g., robot end-effector) is recorded. Each recorded position is augmented with a progress value $\varphi \in [0, 1]$, calculated as the index of that datapoint divided by the total number of datapoints for that demonstration.

The state $x$ is composed of the Cartesian position $\xi$ (2D or 3D) and the progress value $\varphi$, i.e., $x := [\xi, \varphi]$. A demonstration $d_n = \{x_{n,0}, \dots, x_{n,l}\}$, where $n$ denotes the $n$-th demonstration. The set of all training demonstrations is $D^{(0)} = \{d_0, \dots, d_N\}$, in the global (0th) frame.

Before each demonstration, positions and orientations of relevant frames (relative to the global frame) are recorded. For frame $m$, a translation vector $^m t$ and rotation matrix $^m R$ are constructed, transforming any datapoint $x_{n,i}$ from frame 0 to frame $m$ using:

\[
    ^m H =
    \begin{bmatrix} ^m R & 0 \\ 0 & 1
    \end{bmatrix},
\]

\[
    ^m x_{n,i} =
    \begin{bmatrix} ^m t \\ 0
    \end{bmatrix} + ^m H \, x_{n,i}.
\]

The progress variable $\varphi$ is not transformed. This yields $M$ transformed demonstration sets $D^{(1)}, \dots, D^{(M)}$. The policy takes current positions relative to each frame and $\varphi$, outputting $\Delta x$.

\subsection{Alignment of Demonstrations}
To use the progress value effectively, demonstrations are aligned using a modified Dynamic Time Warping (DTW). The assumption is that in each local frame, the closest point in space occurs at the same normalized time.

For each $i$-th demonstration in set $D^{(m)}$, the index $h$ of the closest point $^m x_{i,h}$ to elements of other demonstrations $^m x_{j,:}$ is found:

\[
    ^m A_{ij} = \arg\min_h \| ^m x_{i,h} - ^m x_{j,:} \|,
\]

\[
    ^m B_{ij} = \varphi_{i, A_{ij}}.
\]

The keypoint progress for demonstration $i$ in frame $m$ is:

\[
    P_{im} = \operatorname{median}{(^m B_{i,:})}.
\]

Demonstrations are resampled to align keypoints, yielding aligned sets $D^{(0*)}, \dots, D^{(M*)}$. Fig. 3 illustrates keypoints and alignment for example demonstrations.

\subsection{Training the Local Policies}
Local policies are encoded as dynamical systems:

\[
    \Delta x = f(x),
\]

where $x = [\xi, \varphi]$. Using only position could avoid time misalignment but introduces ambiguity in overlapping trajectories and prevents teaching stopping behaviors. Thus, progress $\varphi$ is included.

The function is modeled as $f(x) \sim \mathcal{GP}(0, k(x, x'))$ with a Matérn kernel and zero mean prior. For large datasets, a variational approximation $q(u) = \mathcal{N}(u | m, S)$ with inducing variables at $Z$ is used, optimized via ELBO.

The predictive distribution at test point $X_*$ is:

\[
    p(f_*) = \mathcal{N}(A m, K(X_*, X_*) + A (S - K(Z, Z)) A^T),
\]

where $A = K(X_*, Z) K(Z, Z)^{-1}$. For multi-output, $p(\Delta x_i) = \mathcal{N}(\mu_i, \sigma_i^2)$, with total variance $\sigma^2(x) = \sum_i \sigma_i^2(x)$.

A risk-minimizing term is added:

\[
    \Delta \xi^{\text{risk}} = -\beta \sigma_i(x) \frac{\nabla_i \sigma^2(x)}{\| \nabla \sigma^2(x) \|},
\]

with $\beta = 2$ for calibration to the prior std, attracting to low-uncertainty regions. This stabilizes dynamics and rejects disturbances, as shown in Fig. 4 for the LASA dataset.

\subsection{Self-Supervised Training of the Frame Relevance GP}
Frame relevance $\alpha \sim \mathcal{GP}(0, k(\varphi, \varphi'))$ is regressed on progress $\varphi$. It is trained self-supervised on held-out demonstrations not used for local GPs.

Local predictions are transformed to global frame using inverse $^m H^{-1}$ and weighted by $\alpha_m$. The combined prediction is Gaussian:

\[
    \theta \mu_i = \sum_{m=1}^M \, ^m \alpha_i \, ^m H^{-1} \, ^m \mu_i (^m x_i),
\]

\[
    \theta \Sigma_i = \sum_{m=1}^M \, ^m \alpha_i \, ^m H^{-1} \, ^m \Sigma_i (^m x_i) \, ^m H^{-T}.
\]

A softmax likelihood ensures weights sum to 1. The likelihood of observed $\Delta x$ is maximized:

\[
    p(\Delta x | f) = \prod_i p(\Delta x_i | f_i),
\]

optimizing to match means while minimizing uncertainty, selecting confident local policies.

The final policy superimposes stabilized local dynamics weighted by relevances for new configurations.

\section{RESULTS \& DISCUSSIONS}

\section{CONCLUSIONS}

We have explored the Task-Parameterized Gaussian Processes (TPGP) framework for learning multi-reference frame skills from demonstrations.
TPGP addresses key challenges in Learning from Demonstration (LfD) by encoding local dynamics relative to task frames using Gaussian Processes and determining frame relevance in a self-supervised manner, without relying on heuristics or manual labeling.
By incorporating progress variables, variance minimization for stability, and probabilistic weighting, TPGP achieves robust generalization to novel configurations, as evidenced by superior performance over Task-Parameterized Gaussian Mixture Models (TP-GMM) in metrics such as goal proximity and Fréchet distance in both simulated multi-frame tasks and real-world robotic re-shelving experiments.

Insights from related works, including interactive imitation learning algorithms like DAgger, HG-DAgger, and CR-DAgger, as well as recent advancements in adaptive motion policies such as keypoint parameterization~\cite{lee2024generalizable} and Elastic Motion Policies~\cite{li2025elastic}, highlight opportunities for enhancing TPGP's adaptability and efficiency in dynamic environments.

A primary limitation of TPGP, shared with TP-GMM, is its dependence on diverse demonstrations for effective generalization.
Similar frame configurations across demonstrations can hinder adaptation to new scenarios.
Future work will focus on integrating TPGP into an incremental learning framework to allow easy incorporation of additional data.
Additionally, incorporating active learning mechanisms, leveraging GP uncertainty quantification, could enable the system to request targeted demonstrations, improving sample efficiency.
Real-robot validations on more complex manipulation tasks, such as compliant interactions or multi-object scenarios, will further assess TPGP's practical utility.
Overall, TPGP represents a promising step toward flexible, user-friendly robot skill acquisition.

\addtolength{\textheight}{-12cm}

\section*{APPENDIX}

\section*{ACKNOWLEDGMENT}

\bibliographystyle{IEEEtran}
\bibliography{references}

\begin{thebibliography}{1}

    \bibitem{calinon2016tutorial}
    S. Calinon, ``A tutorial on task-parameterized movement learning and retrieval'' \emph{Intelligent Service Robotics}, vol. 9, no. 1, pp. 1--29, 2016.

    \bibitem{pignat2017learning}
    E. Pignat and S. Calinon, ``Learning adaptive dressing assistance from human demonstration'' \emph{Robotics and Autonomous Systems}, vol. 93, pp. 61--75, 2017.

    \bibitem{ramirez2024learning}
    M. Ramírez Montero, G. Franzese, J. Kober, and C. Della Santina, ``Learning Multi-Reference Frame Skills from Demonstration with Task-Parameterized Gaussian Processes'' in \emph{2024 IEEE/RSJ International Conference on Intelligent Robots and Systems (IROS)}, 2024.

    \bibitem{ross2011reduction}
    S. Ross, G. J. Gordon, and D. Bagnell, ``A reduction of imitation learning and structured prediction to no-regret online learning'' in \emph{Proceedings of the Fourteenth International Conference on Artificial Intelligence and Statistics}, 2011, pp. 627--635.

    \bibitem{kelly2019hg}
    M. Kelly, C. Sidrane, C. Finn, M. Balakrishna, and Z. Erickson, ``HG-DAgger: Interactive Imitation Learning with Human Experts'' in \emph{2019 International Conference on Robotics and Automation (ICRA)}, 2019.

    \bibitem{balakrishna2019imitation}
    A. Balakrishna, et al., ``On-Policy Robot Imitation Learning from a Converging Supervisor'' in \emph{Conference on Robot Learning (CoRL)}, 2019.

    \bibitem{celemin2022interactive}
    C. Celemin, et al., ``Interactive Imitation Learning in Robotics: A Survey'' \emph{Foundations and Trends in Robotics}, vol. 10, no. 1--2, pp. 1--197, 2022.

    \bibitem{lee2024generalizable}
    S. Lee, et al., ``Generalizable Motion Policies through Keypoint Parameterization and Transportation Maps'' arXiv preprint arXiv:2404.13458, 2024.

    \bibitem{li2025elastic}
    T. Li, S. Sun, S. S. Aditya, and N. Figueroa, ``Elastic Motion Policy: An Adaptive Dynamical System for Robust and Efficient One-Shot Imitation Learning'' arXiv preprint arXiv:2503.08029, 2025.

\end{thebibliography}

\end{document}
